---
title: "Group7_Arumugam_YashwanthBalan_Model2"
author: "YASHWANTH BALAN ARUMUGAM"
date: "29/10/2021"
output:
  html_document: default
  word_document: default
---
Importing and Inspecting Dataset:
  
```{r}
#Importing and Inspecting the Data
data2 <- read.csv("D:/Yash-USA/Northeastern University/Applied Artificial Intelligence/Quarter-1/Predictive Analytics- ALY 6020/Project/Dataset/data2_Yash.csv")
View(data2)
```

Data Cleaning and Preprocessing:
  
```{r}
#Step1-Checking for NA values in the Dataset, if it finds replace with 0
library(dplyr)
data2$SHOOTING <- replace(data2$SHOOTING,is.na(data2$SHOOTING),0)
```

Converting Columns to Numerical Variables:
  
```{r}
#Converting Column 'Shooting': Y-1 and N-0
library(tidyverse)
data2 = data2%>% mutate(SHOOTING = ifelse(SHOOTING == 'Y',1,0),
                        SHOOTING = as.numeric(SHOOTING))
View(data2)
```

```{r}
library(dplyr)
#Converting the Column 'Days_of_week' to its numeric value
data2<-data2%>% mutate(DAY_OF_WEEK_NUM=DAY_OF_WEEK)
data2$DAY_OF_WEEK_NUM <- recode(data2$DAY_OF_WEEK_NUM, 
                                "Sunday"="0",
                                "Monday"="1",
                                "Tuesday"="2",
                                "Wednesday"="3",
                                "Thursday"="4",
                                "Friday"="5",
                                "Saturday"="6")
```

```{r}

#Duplicating the column 'UCR_PART'
data2<-data2%>% mutate(UCR_PART_NUM=UCR_PART)
View(data2)


#Now Converting UCR_PART Values to its Numeric Values
data2$UCR_PART_NUM <- recode(data2$UCR_PART_NUM, 
                             "Part One"="1",
                             "Part Two"="2",
                             "Part Three"="3",
                             "Part Four"="4")
View(data2)
```

Converting the Datatypes:
  
```{r}
#Converting Data Types 
str(data2)
data2$UCR_PART_NUM <- replace(data2$UCR_PART_NUM,is.na(data2$UCR_PART_NUM),1)
transform(data2,UCR_PART_NUM=as.numeric(UCR_PART_NUM))
```

Deleting the Unused Columns:
  
```{r}
data2=data2 %>% select(-DISTRICT)
data2=data2 %>% select(-STREET)
data2=data2 %>% select(-INCIDENT_NUMBER)
data2=data2 %>% select(-OFFENSE_CODE_GROUP)
data2=data2 %>% select(-OCCURRED_ON_DATE)
data2=data2 %>% select(-OFFENSE_CODE)
data2=data2 %>% select(-UCR_PART)
data2=data2 %>% select(-DAY_OF_WEEK)
data2=data2 %>% select(-REPORTING_AREA)
data2=data2 %>% select (-SHOOTING)
data2=data2 %>% select (-Lat)
data2=data2 %>% select (-Long)
data2=data2 %>% select (-Location)

View(data2)
```

Glimpse of New Data:
  
  ```{r}
#Installing 'Tidyverse'
library(tidyverse)
#Having a glimpse of Dataset
glimpse(data2)
```

```{r}
#Converting Data Types:
data2$HOUR<-as.numeric(data2$HOUR)
data2$DAY_OF_WEEK_NUM<-as.numeric(data2$DAY_OF_WEEK_NUM)
data2$UCR_PART_NUM<-as.numeric(data2$UCR_PART_NUM)
str(data2)
```

Summary of New Data:
  
```{r}
summary(data2)
#Cleaning of dataset is done
```

Glipmse of My Data: data2:
  
```{r}
#Import Necessary Libraries
library(dplyr)
glimpse(data2)
```

```{r}
#Verifying continuous variables
continous<-select_if(data2,is.numeric)
summary(continous)
```

```{r}
#compute the top 2 percent percentile
top_one_percent <- quantile(data2$HOUR, .99)
top_one_percent
#Computed that 99% of the crimes happened at 23(11 PM)-Before

```



Model 2- Yashwanth Balan Arumugam Random Forest Model


```{r}
summary(data2)
str(data2)
head(data2)
```

Random Forest Split into Train and Testing Set Training Set : Testing Set = 70 : 30 (random)
```{r}
set.seed(12345)
na.exclude(data2)
data2<-as.data.frame(data2)
train_rf <- sample(nrow(data2), 0.7*nrow(data2), replace = FALSE)
TrainSet_rf <- data2[train_rf,]
TrainSet_rf
```

Testing set of 30%
```{r}
testset_rf <- data2[-train_rf,]
testset_rf
summary(testset_rf)
```

```{r}
# Create a Random Forest model with default parameters
library(randomForest)
model1 <- randomForest(HOUR ~ YEAR+MONTH+DAY_OF_WEEK_NUM+UCR_PART_NUM, data = TrainSet_rf,na.action=na.roughfix, ntree=500, importance=TRUE, do.trace=100)
model1
```

Plotting Model-1
```{r}
plot(model1)
```

```{r}
# Predicting on train set
predTrain_rf <- predict(model1, TrainSet_rf, type = "class")
head(predTrain_rf)
```

```{r}
# Checking classification accuracy
tab_class<-table(predTrain_rf, TrainSet_rf$HOUR)  
head(tab_class)
```

```{r}
# Predicting on Testing set
predtest_rf <- predict(model1, testset_rf, type = "class")
predtest_rf<-na.exclude(predtest_rf)
max(predtest_rf)
#Crimes Mostly to happen between 15 to 16 PM as per the Prediction
```

```{r}
#Accuracy test on Test set
test_error = mean((testset_rf$HOUR - predtest_rf)^2)
print(paste(test_error))

```

Variable Importance plot of Model-1
```{r}
varImpPlot(model1)
```

# Split into Train and Validation sets

# Training Set : Validation Set = 60 : 40 (random)

```{r}
set.seed(12345)
na.exclude(data2)
data2<-as.data.frame(data2)
train_rf_1 <- sample(nrow(data2), 0.6*nrow(data2), replace = FALSE)
TrainSet_rf_1 <- data2[train_rf_1,]
TrainSet_rf_1
```

Testing Set of 40%
```{r}
testset_rf_1 <- data2[-train_rf_1,]
testset_rf_1
summary(testset_rf_1)
```

```{r}
# Create a Random Forest model with default parameters
library(randomForest)
model2 <- randomForest(HOUR ~ YEAR+MONTH+DAY_OF_WEEK_NUM+UCR_PART_NUM, data = TrainSet_rf_1,na.action=na.roughfix, mtry=3, ntree=500, importance=TRUE, do.trace=100)
model2


```

PLotting Model-2
```{r}
plot(model2)
```

```{r}
# Predicting on Testing set
predtest_rf_1 <- predict(model2, testset_rf_1, type = "class")
predtest_rf_1<-na.exclude(predtest_rf_1)
max(predtest_rf_1)
#Crimes Mostly to happen between 15 to 16 PM as per the Prediction
```


```{r}
#Accuracy test on Test set
test_error_1 = mean((testset_rf_1$HOUR - predtest_rf_1)^2)
test_error_1
```
Plotting varibale importance of MOdel-2
```{r}
varImpPlot(model2)
```

# Split into Train and Validation sets

# Training Set : Validation Set = 85:15 (random)

```{r}
set.seed(12345)
na.exclude(data2)
data2<-as.data.frame(data2)
train_rf_2 <- sample(nrow(data2), 0.85*nrow(data2), replace = FALSE)
TrainSet_rf_2 <- data2[train_rf_2,]
TrainSet_rf_2
```

#Creating the test set of 15%
```{r}
testset_rf_2 <- data2[-train_rf_2,]
testset_rf_2
summary(testset_rf_2)
```

```{r}
# Create a Random Forest model with default parameters
library(randomForest)
model3 <- randomForest(HOUR ~., data = TrainSet_rf_2,na.action=na.roughfix, ntree=500, importance=TRUE, do.trace=100)
model3

```

#PLotting Model-3
```{r}
plot(model3)
```

```{r}
# Predicting on Testing set
predtest_rf_2 <- predict(model3, testset_rf_2, type = "class")
predtest_rf_2<-na.exclude(predtest_rf_2)
max(predtest_rf_2)
#Crimes Mostly to happen between 15 to 16 PM as per the Prediction
```

```{r}
#Accuracy test on Test set
test_error_2 = mean((testset_rf_2$HOUR - predtest_rf_2)^2)
test_error_2

```

PLotting the Variable Importance of Model-3
```{r}
varImpPlot(model3)
```

